---
title: "PS3_q1"
author: "Aastha"
date: "3/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE} 
library(tidyverse)
library(readr)
library(ks)
library(kdensity)
library(data.table)
library(fitdistrplus)
library(ggplot2)
library(gridExtra)
library(ggpubr)
library(dvmisc)
library(polywog)
library(splines)
library(freeknotsplines)
library(foreign)
library(MatchIt)
library(dplyr)
library(cem)
library(jtools)
#library(rstatix)
setwd("~/github/empirical-methods/ProblemSets/pset3")
```

Loading DW's version of data.

``` {r, message = FALSE, warning = FALSE}

data_dw <- read.dta("nsw_dw.dta")

```


## Ques 1

Checking for balance across the treatment and control group in the DW sample.

``` {r,  message = FALSE, warning = FALSE}

age <- data_dw$age
treat <- data_dw$treat
black <- data_dw$black
hispanic <- data_dw$hispanic
education <- data_dw$education
married <- data_dw$married
nodegree <- data_dw$nodegree
re74 <- data_dw$re74
re78 <- data_dw$re78
re75 <- data_dw$re75

#mean(age[treat == 1])
#mean(age[treat == 0])

# calculating means by treatment status (control == 0, treatment == 1)

covars<- c('age', 'married', 'education', 'black', 'hispanic', 'nodegree', 're74', 're75')
data_dw %>%
  group_by(treat) %>%
  dplyr::select(one_of(covars)) %>%
  summarise_all(funs(mean(., na.rm = T)))

# Testing for difference in means to check if the data is balanced or not.

lapply(data_dw[,c("married", "age", "black", "nodegree", "hispanic", "education", "re74", "re75")], function(x) t.test(x ~ data_dw$treat, var.equal = TRUE))

```


Yes, the means are balanced (except the 'no degree' variable), the numbers reported for control (treat == 0) and treatment (treat == 1) are the same as table 1 of the paper.


## Ques 2
Estimating the main treatment effet of the training program on 1978 earnings (`re78`)
``` {r, message = FALSE, warning = FALSE}

treatEff_b <- mean(re78[treat == 1]) - mean(re78[treat == 0])
treatEff_b

reg1 <- lm (re78 ~ age + education + black + hispanic + married + nodegree + treat 
            + re74 + re75)

summ(reg1, model.info = FALSE, digits = 3)

reg2 <- lm (re78 ~ age + I(age^2) + education + I(education^2) + nodegree + black + hispanic + re74 + I(re74^2) + re75 + I(re75^2) + married + treat)

summ(reg2, model.info = FALSE, digits = 3)
```


Since the treatment was allocated randomly, the treatment effect without covariates is just the difference of the expected value of the outcome variable under treatment minus expected value of outcome variable under control.


The treatment effect with covariates is a linear regression of the outcome variabe on the treatment dummy variable as well as other control variables. 

Yes, covariates matter. The treatment effect is different when we include covariates (1676.34) versus when the covariates are not included (1794.34). The model specification also matters. If we follow the specification from the paper then the treatment effect is 1622.29. ALl the estimates values of treatment effect are significant at 1% level.

# Ques 3 and 4
Estimating the treatment effect using a different control group from the PSID data. 

``` {r, message = FALSE, warning = FALSE}


# dropping NSW control group

Ndata_dw <- data_dw[which(treat == '1'),]
psid_cont <- read.dta("psid_controls.dta")

dim(Ndata_dw)
dim(psid_cont)

# appending the data with PSID controls
data_dw_psid <- rbind(Ndata_dw, psid_cont)
dim(data_dw_psid)


age <- data_dw_psid$age
treat <- data_dw_psid$treat
black <- data_dw_psid$black
hispanic <- data_dw_psid$hispanic
education <- data_dw_psid$education
married <- data_dw_psid$married
nodegree <- data_dw_psid$nodegree
re74 <- data_dw_psid$re74
re78 <- data_dw_psid$re78
re75 <- data_dw_psid$re75

# Testing for balance between treatment and control
covars_psid <- c('age', 'married', 'education', 'black', 'hispanic', 'nodegree', 're74', 're75')
data_dw_psid %>%
  group_by(treat) %>%
  dplyr::select(one_of(covars_psid)) %>%
  summarise_all(funs(mean(., na.rm = T)))

# testing for difference in means


#lapply(covars_psid, function(v) {
#  t.test(data_dw_psid[, v] ~ data_dw_psid[, 'treat'])
#}) # maybe previous method was nice

lapply(data_dw_psid[,c("married", "age", "black", "nodegree", "hispanic", "education", "re74", "re75")], function(x) t.test(x ~ data_dw_psid$treat, var.equal = TRUE, model.info = FALSE))


#library(rstatix)
#stat.test <- data_dw_psid %>% 
#  t_test(covars_psid ~ treat) %>%
#  add_significance()
#stat.test

```

Notice that the samples are not balanced across observables in the control versus treatment group. The difference between the mean values of the covariates across the treatment and control group are significantly different, as can be seen by the p-values in the results above. Therefore, after changing the control group to the PSID sample, the treatment and the control groups are no more balanced in the observables. 

#Ques 5

Constructing a propensity score for the treatment.

``` {r, message = FALSE, warning = FALSE}


# estimate the propensity score by running a logit model


m_ps <- glm(treat ~ age + education + nodegree + black + hispanic + re74 + re75 + married , family = binomial(), data = data_dw_psid)
summary(m_ps)


# Using this model, we can now calculate the propensity score for observation. It is simply the sample's predicted probability of being Treated, given the estimates from the logit model. Below, I calculate this propensity score using predict() and create a dataframe that has the propensity score as well as the actual treatment status.

prs_df <- data.frame(pr_score = predict(m_ps, type = "response"),
                     treat = m_ps$model$treat)
head(prs_df)

```

Plotting the propensity score.

``` {r, message = FALSE, warning = FALSE}

labs <- paste("Treatment Status:", c("Treated", "Control"))
prs_df %>%
  mutate(treat = ifelse(treat == 1, labs[1], labs[2])) %>%
  ggplot(aes(x = pr_score)) +
  geom_histogram(color = "white") +
  facet_wrap(~treat) +
  xlab("Probability of getting treated") + xlim(0,1) + ylim(0,80) +
  theme_bw()



```

Replotting using the specification used in footnote of table 3:

``` {r, echo=FALSE, message = FALSE, warning = FALSE}
m_ps3 <- glm(treat ~ age + I(age^2) + education + I(education^2) + nodegree + black + hispanic + re74 + I(re74^2) + re75 + I(re75^2) + married , family = binomial(), data = data_dw_psid)
summary(m_ps3)




prs_df3 <- data.frame(pr_score3 = predict(m_ps3, type = "response"),
                     treat = m_ps3$model$treat)
head(prs_df3)

labs <- paste("Treatment Status:", c("Treated", "Control"))
prs_df3 %>%
  mutate(treat = ifelse(treat == 1, labs[1], labs[2])) %>%
  ggplot(aes(x = pr_score3)) +
  geom_histogram(color = "white") +
  facet_wrap(~treat) +
  xlab("Probability of getting treated") + xlim(0,1) + ylim(0,80) +
  theme_bw()



```

# Ques 6 Re-estimating the treatment effect after matching on the estimated propensity score, using nearest-neighbor matching (with replacement)

The method we use below is to find pairs of observations that have very similar propensity scores, but that differ in their treatment status. We use the package MatchIt for this. This package estimates the propensity score in the background and then matches observations based on the method of choice (“nearest” in this case)

``` {r, message = FALSE, warning = FALSE}


data_dw_psid_nomiss <- data_dw_psid %>%  # MatchIt does not allow missing values
  dplyr::select(re78, treat, one_of(covars_psid)) %>%
  na.omit()


mod_match <- matchit(treat ~ age + I(age^2) + education + I(education^2) + nodegree + black + hispanic + re74 + I(re74^2) + re75 + I(re75^2) + married,
                     method = "nearest", data = data_dw_psid_nomiss)

summary(mod_match)

plot(mod_match)

```


Creating a data frame containing only the matched observations:

``` {r, message = FALSE, warning = FALSE}
# To create a dataframe containing only the matched observations, use the match.data() function:

dta_m <- match.data(mod_match)
dim(dta_m)

```

Estimating treatment effects:

``` {r, message = FALSE, warning = FALSE}
with(dta_m, t.test(re78 ~ treat))

# or use OLS with or without controls 
lm_treat1 <- lm(re78 ~ treat, data = dta_m)
summ(lm_treat1, digits = 2, model.info = FALSE)


lm_treat2 <- lm(re78 ~ age + I(age^2) + education + I(education^2) + nodegree + black + hispanic + re74 + I(re74^2) + re75 + I(re75^2) + married + treat, data = dta_m)
summ(lm_treat2, digits = 2, model.info = FALSE)
```


#Ques 7 estimating the treatment effect using coarsened exact matching(CEM)
Using the CEM package below:

``` {r, message = FALSE, warning = FALSE}
library(cem)

data <- data.frame(na.omit(data_dw_psid))

imbalance(group=data$treat, data=data[covars_psid]) # common support is 1.4%

mat <- cem(treatment = "treat", data = data, drop = "re78",keep.all=TRUE)
mat

est <- att(mat, re78 ~ treat + age + I(age^2) + education + I(education^2) + nodegree + black + hispanic + re74 + I(re74^2) + re75 + I(re75^2) + married, data = data)
est

est2 <- att(mat, re78 ~ treat + age + I(age^2) + education + I(education^2) + nodegree + black + hispanic + re74 + I(re74^2) + re75 + I(re75^2) + married, data = data, model = "forest")
est2


plot(est2, mat, data, vars=c("education", "age", "re74", "re75", "black", "hispanic", "nodegree", "married"))
```
Coarsened exact matching seems to perfrom better than the previous matching method. Random forest model estimate including the covariates gives a slightly "better" treatment effect. The treatment effect estimate is 923.54.