---
title: "PS3_q2"
author: "Aastha"
date: "3/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r, echo=FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(readr)
library(ks)
library(kdensity)
library(data.table)
library(fitdistrplus)
library(ggplot2)
library(gridExtra)
library(ggpubr)
library(dvmisc)
library(polywog)
library(splines)
library(freeknotsplines)
library(foreign)
library(MatchIt)
library(dplyr)
library(cem)
setwd("~/github/empirical-methods/ProblemSets/pset3")
```

``` {r, message = FALSE, warning = FALSE}
data <- load("yelp.Rdata")

#rev <- data$logrev

```



In the following steps, I am loading the data 10 times and using a different discontinuity each time starting with 0.25, 0.75, 1.25 and so on.. For example, there is a discontinuity at 0.25, if the score is 0.24 for example then the stars given are rounded to 0, whereas if the score is 0.26 then it is rounded-off to 0.5 stars. I estimate the effect of these 10 different discontinuities on the dependendent variable : log of revenue. The effect of different discontinuities are given below.
``` {r, message = FALSE, warning = FALSE}
data_1 <- yelp %>% subset(score<0.75) %>% mutate(treat = ifelse(stars == 0.5 , 1, 0))
data_2 <- yelp %>% subset(score>=0.25 & score<1.25) %>% mutate(treat = ifelse(stars == 1 , 1, 0))
data_3 <- yelp %>% subset(score>=0.75 & score<1.75) %>% mutate(treat = ifelse(stars == 1.5 , 1, 0))
data_4  <- yelp %>% subset(score>=1.25 & score<2.25) %>% mutate(treat = ifelse(stars == 2 , 1, 0))
data_5  <- yelp %>% subset(score>=1.75 & score<2.75) %>% mutate(treat = ifelse(stars == 2.5 , 1, 0))
data_6  <- yelp %>% subset(score>=2.25 & score<3.25) %>% mutate(treat = ifelse(stars == 3 , 1, 0))
data_7  <- yelp %>% subset(score>=2.75 & score<3.75) %>% mutate(treat = ifelse(stars == 3.5 , 1, 0))
data_8  <- yelp %>% subset(score>=3.25 & score<4.25) %>% mutate(treat = ifelse(stars == 4 , 1, 0))
data_9  <- yelp %>% subset(score>=3.75 & score<4.75) %>% mutate(treat = ifelse(stars == 4.5 , 1, 0))
data_10 <- yelp %>% subset(score>=4.25) %>% mutate(treat = ifelse(stars == 5 , 1, 0))

rd_eff <- data.frame(matrix(NA, nrow = 10, ncol = 1))

for (i in 1:10){
   reg <- lm(logrev ~ treat + score, data = get(paste0("data_" , i)))
  rd_eff[i,1] <- summary(reg)$coefficients[2,1]
}


```

 Reporting the average treatment effect for 10 different treatments below

``` {r, message = FALSE, warning = FALSE}
colnames(rd_eff) <- 'treatment effect for 10 discontinuities'
print(rd_eff)
```
now finding the average treatment effect across the 10 discontinuities


``` {r, message = FALSE, warning = FALSE}
avg_eff <- mean(rd_eff$`treatment effect for 10 discontinuities`)
avg_eff
```

Average treatment effect across 10 discontinuitie is positive and significant, of the magnitude 1.375
``` {r, message = FALSE, warning = FALSE}

```



``` {r, message = FALSE, warning = FALSE}

```




``` {r, echo=FALSE, message = FALSE, warning = FALSE}

```



``` {r, echo=FALSE, message = FALSE, warning = FALSE}

```